---
title: "Exercise 3.2"
author: "Joe Barry"
date: "May 6, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library('tswge')
data("dowjones2014")
```

Now that the data is loaded, we can use the Akaike Information Criterion (AIC) to produce our best model fit.

```{r}
aic.wge(dowjones2014, p=0:12, q=0:4, type="aicc")
aic.wge(dowjones2014, p=0:12, q=0:4, type="aic")
aic.wge(dowjones2014, p=0:12, q=0:4, type="bic")
```

For each model identifier criterion, the value of p chosen was $p=1$.

```{r}
plot.ts(ts(dowjones2014),
        main = 'Daily Dow Jones Industrial Average, 2014',
        ylab = 'Stock Price', 
        xlab = 'Day')
period.wge(dowjones2014)
acf(dowjones2014, lag.max = 100)
```

Based on the periodogram and ACF, this data is most similar to a random walk realization, in that the value at any given point in time is some normally distributed distance from the previous value. This is reflected in the lack of a dominant frequency beyond the first measured in the periodogram, and an autocorrelation which slowly falls to zero.



```{r}
yule <- est.ar.wge(dowjones2014, p=1, factor = FALSE, type='yw')
burg <- est.ar.wge(dowjones2014, p=1, factor = FALSE, type='burg')
mle <- est.ar.wge(dowjones2014, p=1, factor = FALSE, type='mle')

```

```{r}
paste('YW phi_1:', yule$phi)
paste('Burg phi_1:', burg$phi)
paste('MLE phi_1:', mle$phi)

paste('YW variance:', yule$avar)
paste('Burg variance:',burg$avar)
paste('MLE variance:',mle$avar)
```

So our results are as follows:

|  | $\phi_1$ | $\sigma^2$ | 
| ----------: | ----------: | ---------------: | 
| Yule-Walker | .970835 | 12955 |
| Burg | .97853 | 12907 |
| Maximum Likelihood| .98159 | 12898 |


From this data, we can assume that the Maximum Likelihood Estimator is our best model, as it has the lowest residual variance.







