---
title: "Exercise 1.2"
author: "Kelton Berry"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Question 1:
To begin, we must import both the library and data we will be working with.

```{r}
library("tswge")
data(ss08)
```

Now we can plot this data first with thebase plot in R.

```{r}
plot(ss08, type='l')
```

We then can plot the data with plotts.wge function provided by our textbook
```{r}
plotts.wge(ss08)
```

#Question 2

Here, we can find the mean of our dataset. 
```{r}
mean(ss08)
```

#Question 3 

Here we will load the hadley dataset and plot the sample autocorrelation and sample autocovariance using the acf function. 

```{r}
data("hadley")
```

```{r}
acf(hadley, type = "correlation")
```

```{r}
acf(hadley, type = "covariance")
```

#Question 4

Here, I will write the equations for ^γ0, ^γ1, ^ρ0 and ^ρ1 and compute them using R.

First, covariance:
$$\hat{\gamma_{0}}= \frac{1}{n}\sum_{t=1}^{n}{(X_t-\mu)}* ({X}_t-\mu)$$

$$\hat \gamma_1 = \frac{1}{n} \sum_{t=1}^{n-|1|} (X_t - \mu)*(X_{(t + 1)} - \mu)$$
And now, autocorrelation:

$$\hat \rho_0 = \frac{\hat \gamma_0}{\hat \gamma_0}$$
$$\hat \rho_1 = \frac{\hat \gamma_1}{\hat \gamma_0}$$


Now, we will calculate autocovariance and autocorrelation in R:
```{r}
acf(hadley, lag.max = 1, plot = FALSE)
```
```{r}
acf(hadley, lag.max = 1, type = 'covariance', plot = FALSE)
```

## Question 5: 
5.Plot the time series and the periodogram for Figure 1.10a of Woodward et al. (fig1.10a).

Here we plot the time series:
```{r}
data("fig1.10a")
plotts.wge(fig1.10a)
```
And here we plot the Periodogram
```{r}
periodogram <- period.wge(fig1.10a)
```
Now we need to find the three dominate frequencies, we are apparent from the plot:
```{r}
periodogram <- period.wge(fig1.10a)$pgram
f_1 <- period.wge(fig1.10a)$freq[which.max(periodogram)]* 10
f_1
```




```{r}
periodogram <- periodogram[-which.max(periodogram)]
f_2 <-period.wge(fig1.10a, dbcalc = TRUE, plot = FALSE)$freq[which.max(periodogram)]* 10
f_2
```

```{r}
periodogram <- periodogram[-which.max(periodogram)]
f_3<-period.wge(fig1.10a, dbcalc = TRUE, plot = FALSE)$freq[which.max(periodogram)]* 10
f_3
```

The three dominate frequencies are .15, 0.03 and 0.01

##Question 6
Replace the frequencies you found in the equation above and recreate using R the Figure 1.10a using values for t ∈ (0, 100)

```{r}
t <- 0:1000/10
X_t <- cos(2*pi*t*0.025) + 1.5*cos(2*pi*t*(0.1)+1)+2*cos(2*pi*t*(0.15)+2.5)
plot.ts(X_t)
```

