---
title: "Exercise 1.2"
author: "Asha M Silva" 
date: "2/25/2021"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tswge")
```


# Question 1

```{r}

# load in the dataset 
data("ss08")
# Base plot in R
plot(ss08, type = "l")
# Plot using function provided by the textbook
plotts.wge(ss08)
```

# Question 2
Calculate the average of this time series realization.

```{r}
mean(ss08)
# mean = 52.24385
```

# Question 3

Load datset
```{r}
data(hadley)
```

sample autocorrelations and autocovariances
```{r}
#autocorrelations
acf(hadley, type = "correlation")

#autocovariance
acf(hadley, type = "covariance")
```

# Question 4
Write the equations for ^γ0, ^γ1, ^ρ0 and ^ρ1 and compute them using R (note the hats, ^, are to represent the estimators).

## Autocovariance:
$$\hat \gamma_0 = \frac{1}{n} \sum_{t=1}^{n} (X_t - \mu) (X_{t} - \mu) $$
$$\hat \gamma_1 = \frac{1}{n} \sum_{t=1}^{n-|1|} (X_t - \mu) (X_{t + |1|} - \mu)$$
### Autocovariance Answers
```{r}
acf(hadley, lag.max = 1, type="covariance", plot = FALSE)
```

$$\hat \gamma_0 = 0.0667$$ 
$$\hat \gamma_1 = 0.0588$$

## Autocorrelation:
$$\hat \rho_0 = \frac{\hat \gamma_o}{\hat \gamma_0} $$
$$\hat \rho_1 = \frac{\hat \gamma_1}{\hat \gamma_0}$$
### Autocorrelation answers
```{r}
acf(hadley, lag.max = 1, plot = FALSE)
```

$$\hat \rho_0 = 1.000$$ 

$$\hat \rho_1 = 0.882$$

# Question 5

Plot the time series and the periodogram for Figure 1.10a of Woodward et al. (fig1.10a). Note that 
the sampling rate for the x-axis is in 1/10ths. What are the three dominant frequencies of this 
dataset?

Load the dataset:
```{r}
data(fig1.10a)
```

Plot the data:
```{r}
plot.ts(fig1.10a)
```

Plot the periodogram:
```{r}
periodogram <- period.wge(fig1.10a, dbcalc = TRUE, plot = TRUE)
```

What are the three dominant frequencies?
The dominant frequencies are 15, 10, and 3 before scaling.
Taking the scale into account - the dominance frequencies are 0.15, 0.10, and 0.03
```{r}
which.max(periodogram$pgram)
# 15
which.max(periodogram$pgram[1:14])
#10
which.max(periodogram$pgram[1:9])
#3
print("The dominant frequencies are 0.15, 0.10, and 0.03.")
```

# Question 6
The time series in Figure 1.10a (Woodward et al.) can be expressed as the following:
X(t) = cos(2π t f1) + 1.5 cos(2π t f2 + 1) + 2 cos(2 π t f3 + 2.5)

where f1, f2, and f3 are the three dominant frequencies (in order from lowest to highest) found in the previous question. Replace the frequencies you found in the equation above and recreate using R the Figure 1.10a using values for t ∈ (0, 100) (e.g., x <- 0:1000/10).

Originally:
$$X(t) = \cos(2\pi t f_1) + 1.5 \cos(2\pi t f_2 + 1) + 2 \cos(2 \pi t f_3 + 2.5)$$
Given our dominant frequencies and $t \in (0, 100)$:
$$X(t) = \cos(2\pi t (.03)) + 1.5 \cos(2\pi t (.1) + 1) + 2 \cos(2 \pi t (.15) + 2.5)$$

#### Plotting $X(t)$
```{r}
f1 <- 0.03
f2 <- 0.10
f3 <- 0.15
t <- 0:1000/10
Xt <- (cos(2*pi*t*f1) + (1.5*cos((2*pi*f2*t)+1)) + (2*cos((2*pi*f3*t)+2.5)))
plotts.wge(Xt)
plot.ts(Xt)
```


savehistory(file = "exercise_1.2_ashamsilva.Rhistory")

