---
title: "johnhennin_b1_1.2"
author: "John Hennin"
date: "2021-02-20"
output: html_document
---

## Question 1

#### Download twsge package and ss08 data
```{r}
library("tswge")
data(ss08)
```

#### Create plot using base plot for R
```{r}
plot(ss08, type="l")
```

#### Create plot using plotts function from textbook
```{r}
plotts.wge(ss08)
```


## Question 2

#### Calculate mean of ss08 data
```{r}
mean(ss08)
```

## Question 3

#### Load hadley data
```{r}
data(hadley)
```

#### Plotting the autocorrelation
```{r}
acf(hadley, type=c("correlation"), lag.max = 80)
```

#### Plotting the autocovariance
```{r}
acf(hadley, type=c("covariance"), lag.max = 80)
```

## Question 4

The estimate of the autocovariance equation is:
$$ \hat \gamma_h = \frac{1}{n} \sum_{t=1}^{n-|h|} (X_t - \mu) (X_{t + |h|} - \mu) $$

Therefore:

$$\hat\gamma_0 = \frac{1}{n} \sum_{t=1}^{n} (X_t - \mu)^2 = {\sigma^2}$$

$$\hat\gamma_1 = \frac{1}{n} \sum_{t=1}^{n-1} (X_t - \mu) (X_{t + 1} - \mu) $$

#### Calculating $\hat\gamma_0$
```{r}
acf(hadley, type=c("covariance"), lag.max = 80, plot = FALSE)$acf[1]
```

#### Calculating $\hat\gamma_1$
```{r}
acf(hadley, type=c("covariance"), lag.max = 80, plot = FALSE)$acf[2]
```

The estimate of the autocorrelation equation is:
$$\hat \rho_h = \frac{\hat \gamma_h}{\hat \gamma_0}$$

Therefore:

$$ \hat\rho_0 = \frac{\hat\gamma_0}{\hat \gamma_0} =  1$$

$$ \hat\rho_1 = \frac{\hat\gamma_1}{\hat \gamma_0} $$

#### Calculating $\hat\rho_1$
```{r}
acf(hadley, type=c("correlation"), lag.max = 80, plot = FALSE)$acf[2]
```

## Question 5

#### Plotting the time series
```{r}
data("fig1.10a")
plotts.wge(fig1.10a)
```

#### Plotting the periodogram
```{r}
period.fig1.10a <- period.wge(fig1.10a)
```

#### Finding the three dominant frequencies of fig1.10a
```{r}
# Note the frequencies are multiplied by 10 to account for the sampling rate

# Saving periodogram values
pgrams <- period.wge(fig1.10a, plot = FALSE)$pgram

# Finding maximum periodogram value and cooresponding frequency
f.3 <- period.wge(fig1.10a, plot = FALSE)$freq[which.max(pgrams)]*10

# Removing previous maximum from vector of periodogram values
pgrams <- pgrams[-which.max(pgrams)]

# Finding new maximum periodogram value and cooresponding frequency
f.2 <- period.wge(fig1.10a, plot = FALSE)$freq[which.max(pgrams)]*10

# Removing previous maximum from vector of periodogram values
pgrams <- pgrams[-which.max(pgrams)]

# Finding new maximum periodogram value and cooresponding frequency
f.1 <- period.wge(fig1.10a, plot = FALSE)$freq[which.max(pgrams)]*10

cat("The dominant frequencies are (in descending order):", f.3, f.2, f.1)
```

## Question 6

Originally:
$$X(t) = \cos(2\pi t f_1) + 1.5 \cos(2\pi t f_2 + 1) + 2 \cos(2 \pi t f_3 + 2.5)$$
Given our dominant frequencies and $t \in (0, 100)$:
$$X(t) = \cos(2\pi t (.03)) + 1.5 \cos(2\pi t (.1) + 1) + 2 \cos(2 \pi t (.15) + 2.5)$$

#### Plotting $X(t)$
```{r}
t <- 1:1000/10
X.t <- (cos(2*pi*t*f.1) + (1.5*cos((2*pi*f.2*t)+1)) + (2*cos((2*pi*f.3*t)+2.5)))

plotts.wge(X.t)
```

