---
title: "jpdas18_b1_1.2"
name: "Jacinta Das"
Date: "Feb 26, 2020"
output: html_document
---

``` {r}
install.packages("tswge")
library("tswge")
```
### Question 1: Load the dataset ss08 and visualize it. Try the base plot and the plotts.wge function. 
```{r}
data(ss08)
plot(ss08, type="l")
plotts.wge(ss08)
``` 
### Question 2: Calculate the average of this time series realization using the mean function.
```{r}
cat("Mean;", mean(ss08))
``` 
### Question 3: Load the dataset hadley and plot the sample autocorrelations and sample autocovariance using the acf function
```{r}
data(hadley)
acf(hadley, type="covariance")
acf(hadley, type="correlation")
``` 
### Question 4: Write equations for sample autocovariance and autocorrelation at lag=0 and lag=1.
$$\hat\gamma_0 = \frac{1}{n}\sum_{t=1}^{n}(X_t-\mu)^2 = \sigma^2 $$

$$\hat\gamma_1 = \frac{1}{n}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1} - \mu)$$
$$\hat\rho_0 =\frac{\hat\gamma_0}{\hat\gamma_0}  =1$$
$$\hat\rho_1 = \frac{\hat\gamma_1}{\hat\gamma_0}$$
### Compute autocovariance and autocorrelation at lags 0 and 1. 
```{r}
acf(hadley, type="covariance", lag.max=0, plot = FALSE)
acf(hadley, type="correlation", lag.max=0, plot = FALSE)
acf(hadley, type="covariance", lag.max=1, plot = FALSE)
acf(hadley, type="correlation", lag.max=1, plot = FALSE)
```

### Question 5: Plot the time series and the periodogram from Figure 1.10a of Woodward et al. (fig 1.10a). What are the three dominant frequencies of the dataset? I multiplied these by 10 because the sampling rate for the x-axis is in 1/10ths.
```{r}
data(fig1.10a)
plot(1:1000/10, fig1.10a, type='l')
spectrum(fig1.10a)
p.f110a <- period.wge(fig1.10a, dbcalc = TRUE, plot = TRUE)
which.max(p.f110a$pgram)
cat("Frequency 1 is: " , p.f110a$freq[15]*10)
which.max(p.f110a$pgram[0:14])
cat("Frequency 2 is: ", p.f110a$freq[10]*10)
which.max(p.f110a$pgram[0:9])
cat("Frequency 3 is: ", p.f110a$freq[3]*10)
```

### Question 6: Replace the frequencies you found in the equation above and recreate using R the Figure 1.10a using values for $t \in (0,100)$. Frequencies are in order from lowest to highest, so I switched the order of the frequencies from my original calculations.

```{r}
t <- 1:1000/10
Xt <- cos(2*pi*t*.03) + (1.5*cos((2*pi*t*.1)+1)) + 2*cos((2*pi*t*.15)+ 2.5)
plotts.wge(Xt)

```
