---
title: "Exercise 1.2"
author: "Monica Alicea" 
output: html_document
---

#1.Load the dataset ss08 and visualize it. Try both the base plot in R and the plotts.wge function provided by the textbook.
```{r}
library(tswge)
data("ss08")
plot(ss08, type = 'l') #base plot
plotts.wge(ss08) #plotts.wge
```

#2.Calculate the average of this time series realization using the mean function.
```{r}
mean(ss08)
```
The mean is 52.24385

#3.Load the dataset hadley and plot the sample autocorrelations and sample autocovariances using the acf function

```{r}
data(hadley) 
acf(hadley, type= "correlation") #plot sample autocorrelation
acf(hadley, type= "covariance") #plot sample autocovariance
```

#4.Write the equations for ^γ0, ^γ1, ^ρ0 and ^ρ1 and compute them using R (note the hats, ^, are to represent the estimators).
$$
\hat{\gamma_{0}}= \frac{1}{n}\sum_{t=1}^{n}{(X_t-\mu)}- ({X}_t-\mu)
$$
$$
\hat{\gamma_{1}}= \frac{1}{n}\sum_{t=1}^{n-|1|}{(X_t-\mu- ({X}_{t+|1|}-\mu)
$$
$$
\hat{\rho_0}= \frac{\hat\gamma_0}{\hat\gamma_0}=1
$$
$$
\hat{\rho_1} =\frac{\hat\gamma_1}{\hat\gamma_0}
$$

```{r}
acf(hadley, lag.max = 1, plot = FALSE)
```
```{r}
acf(hadley, lag.max = 1, type = 'covariance', plot = FALSE)
```
The autocorrelations are (0,1) and (1,0.882). The autocovariances are (0,0.0667) and (1,0.0588)

#5.Plot the time series and the periodogram for Figure 1.10a of Woodward et al. (fig1.10a). Note that the sampling rate for the x-axis is in 1/10ths. What are the three dominant frequencies of this dataset?
```{r}
data("fig1.10a")
plotts.wge(fig1.10a)
periodogram <- period.wge(fig1.10a, plot = TRUE) #plot periodogram
```
```{r}
#first frequency
freq_1 <- period.wge(fig1.10a, plot = FALSE)$freq[which.max(period.wge(fig1.10a, plot = FALSE)$pgram[0:9])]
freq_1 <- freq_1*10 #to account for the sampling rate

#second frequency 
freq_2 <- period.wge(fig1.10a, plot = FALSE)$freq[which.max(period.wge(fig1.10a, plot = FALSE)$pgram[0:14])] 
freq_2 <- freq_2*10 

#third frequency 
freq_3 <- period.wge(fig1.10a, plot = FALSE)$freq[which.max(period.wge(fig1.10a, plot = FALSE)$pgram)]
freq_3 <- freq_3*10 
```
The three frequencies are 0.03, 0.1, 0.15

#6.The time series in Figure 1.10a (Woodward et al.) can be expressed as the following: X(t) = cos(2π t f1) + 1.5 cos(2π t f2 + 1) + 2 cos(2 π t f3 + 2.5) where f1, f2, and f3 are the three dominant frequencies (in order from lowest to highest) found in the previous question. Replace the frequencies you found in the equation above and recreate using R the Figure 1.10a using values for t ∈ (0, 100) (e.g., x <- 0:1000/10).

```{r}
t <- 0:1000/10
Xt <- cos(2*pi*t*freq_1) + 1.5*cos(2*pi*t+1*freq_2) + 2*cos(2*pi*t + 2.5*freq_3)
plot.ts(Xt)
```