---
Name: Kimya Shirazi
Course: Applied Time Series Analysis
Exercise 1.2
title: "zkshirazi_b1.RMD"
output: html_document
---
## Question 1 
First, Load the tswge library
```{r}
library('tswge')
```

Then, plot the data using the built-in and tswge functions
```{r}
data(ss08)
plot(ss08, type = 'l')
plotts.wge(ss08)
```

## Question 2 
To find the mean using the mean() function, 
```{r}
mean(ss08)
```

## Question 3
Load the hadley dataset
```{r}
data(hadley)
```

Plot the sample autocorrelations and autocovariance
```{r}
acf(hadley, type = 'correlation')
acf(hadley, type = 'covariance')
```

## Question 4
$$\hat \gamma_0 = \frac{1}{n} \sum_{t=1}^{n} (X_t - \mu) (X_{t+1} - \mu)$$

$$\hat \gamma_1 = \frac{1}{n} \sum_{t=1}^{n-1} (X_t - \mu) (X_{t} - \mu)$$ 
$$\hat \rho_0 = \frac{\hat \gamma_0}{\hat \gamma_0}$$
$$\hat \rho_1 = \frac{\hat \gamma_1}{\hat \gamma_0}$$
To calculate them, in order:

```{r}
acf(hadley, lag_max = 1, type = 'correlation', plot = FALSE)
acf(hadley, lag_max = 1, type = 'covariance', plot = FALSE)
```

## Question 5 
plot the time series and periodogram:
```{r}
data(fig1.10a)
plotts.sample.wge(fig1.10a)
```

The plot shows that the three dominant frequencies all lie within a very small window. 
To calculate the exact values:
```{r}
prdgram <- period.wge(fig1.10a, dbcalc = TRUE, plot = FALSE)$pgram
period.wge(fig1.10a, dbcalc = TRUE, plot = FALSE)$freq[which.max(prdgram)]* 10
#0.15 
prdgram <- prdgram[-which.max(prdgram)]
period.wge(fig1.10a, dbcalc = TRUE, plot = FALSE)$freq[which.max(prdgram)]* 10
#0.1
prdgram <- prdgram[-which.max(prdgram)]
period.wge(fig1.10a, dbcalc = TRUE, plot = FALSE)$freq[which.max(prdgram)]* 10
#0.03
```
The 3 dominant frequencies are 0.15, 0.1, and 0.03

## Question 6
```{r}
f1 <- 0.03
f2 <- 0.1
f3 <- 0.15
t <- 0:1000/10
Xt <- cos(2*pi*t*f1)+(1.5*cos((2*pi*t*f2)+1)) + (2*cos((2*pi*t*f3)+2.5))
plotts.wge(Xt)
```