---
title: "Exercise 3.1"
author: "Joe Barry"
date: "May 4, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 3.1.1

```{r}
library('tswge')
```

First, we read in the data from our .txt file.

```{r}
complete.url <- 'https://www.esrl.noaa.gov/gmd/webdata/ccgg/trends/co2/co2_mm_mlo.txt'
co2.df <- read.table(complete.url, col.names = c('year','month','decimal date','monthly average','deseasonalized','ndays','stdays', 'uncertainty'))
```

```{r}
head(co2.df)
```

Now, to plot this as a time series, we have to convert both the averages and the seasonalized averages to time series with monthly periods.

```{r}
deseason <- ts(co2.df$deseasonalized, frequency = 12, start = 1958.203)
avg <- ts(co2.df$monthly.average, frequency = 12, start = 1958.203)
ts.plot(deseason, avg, #start is based on first decimal date entry
        xlab = "Year",
        ylab = "parts per million (ppm)",
        main = "Atmospheric CO2 at Mauna Loa Oservatory",
        col = c('black','red'))
```

This is our complete graph! Next, we want to derive the deseasonalized data ourselves. This can be done with the stl() function.

```{r}
decomp <- stl(avg, 'periodic')
decomp.df <- decomp$time.series
plot.ts(decomp.df[,2],
        xlab = 'Year',
        ylab = 'parts per million (ppm)',
        main = 'Deseasonalized CO2 Levels, 1958-2021')
```


# Exercise 3.1.2

The formula from class which we used for the autocorrelation of $AR(q)$ models was as follows: 
$$
\begin{align}
\rho_k &= \frac{-\theta_k + \sum_{j=1}^{q-k} \theta_j \theta_{j+k}}{1 + \sum_{j=1}^{q} \theta^2_j}, \, &k=1,2,..,q \\
&= 0,\, &k > q
\end{align}
$$


From this and the definition of autocorrelation, we can already determine that $\rho_0 = 1$ (as is the case for any time series, at lag 0 the realization is perfectly correlated), and that $\rho_3 = 0$ (We only have an MA(2) model, so for lags beyond $k = 2$ the autocorrelation is 0). 

The equation we are given is $X_t = a_t + 0.2 a_{t-1} - 0.48 a_{t-2}$. 

So, $\theta_1 = .2$ and $\theta_2 = -.48$

Now, we can use the formula to calculate $\rho_1$:

```{r}
theta_1 <- .2
theta_2 <- -.48
rho_1 <- (-1*theta_1 + theta_1*theta_2)/(1 + (theta_1^2) + (theta_2^2))
print(rho_1)
```

So, $\rho_1 = -.2330$.

Next, we will calculate $\rho_2$:

```{r}
rho_2 <- (-1*theta_2)/(1 + (theta_1^2) + (theta_2^2))
print(rho_2)
```

For our last value, $\rho_2 = .3778$.


# Exercise 3.1.3

For an $AR(1)$ model, the formula given for autocorrelation is: 
$$
\rho_k = {\phi_1}^{|k|}
$$

With $\phi_1 = -.7$ given, all we need to do is take this formula from $k=1$ to $k=10$.

```{r}
k <- c(1:10)
phi_1 <- -.7
rho_1to10 <- phi_1^(k)
rho_1to10
```

So, here is our answer:

| $k$         | $\rho_k$         | 
| ----------: | ---------------: | 
| ----------- | ---------------- |
| 1 | -.7 |
| 2| .49 |
| 3 | -.343 |
| 4 | .2401 |
| 5 | -.1681 |
| 6 | .1176 |
| 7 | -.0824 |
| 8 | .0576 |
| 9 | -.0404 |
| 10 | .0282 |




