---
title: "Exercise 3"
author: "Peter Woo"
date: "5/3/2021"
output: html_document
---

# 3.1.1
PROBLEM STATEMENT:
Reproduce the figure above of atmospheric CO2 concentrations (both seasonal 
and deseasonalized). Can you create the deseasonalized time series using R
modeling tools? Write down your thoughts on the methodology (i.e., prepare a 
pseudo code version) and begin implementing code that creates the 
deseasonalized CO2 trend line.

## Part 0 - Getting the data

```{r}
data <- read.table('https://www.esrl.noaa.gov/gmd/webdata/ccgg/trends/co2/co2_mm_mlo.txt')
names(data) <- c("year", "month", "decidate","moco2", "deseason", "nodays", "stdays", "uncertainty" )

```

## Part 1 - Plot reproduction

```{r}
options(repr.plot.width=12, repr.plot.height=4, repr.plot.res = 125)
plot(data$decidate, 
     data$moco2,
     type = 'l',
     col = 'red',
     lwd = 2,
     main = "Atmospheric CO2 at Mauna Loa Observatory",
     xlab = "Year",
     ylab = "Parts Per Million (ppm)"
     )
lines(data$decidate, data$deseason)
```

## Part 2 - Deseasonalize the monthly time series 

```{r}
co2.decomp <- stl(co2, 'periodic')
trend <- co2.decomp$time.series[, 'trend']

options(repr.plot.width=12, repr.plot.height=10, repr.plot.res = 125)

plot(data$decidate, 
     data$moco2,
     type = 'l',
     col = 'red',
     lwd = 2,
     main = "Atmospheric CO2 at Mauna Loa Observatory",
     xlab = "Year",
     ylab = "Parts Per Million (ppm)"
)
lines(trend)

# shorten the length of moco2 to fit the trend line

shortened <- data$moco2[1:(length(trend))]

plot(data$decidate[1:(length(trend))], 
     shortened,
     type = 'l',
     col = 'red',
     lwd = 2,
     main = "Atmospheric CO2 at Mauna Loa Observatory",
     xlab = "Year",
     ylab = "Parts Per Million (ppm)"
)
lines(trend)

# subtract trend from moco2 for deseasonalized plot

deseasonalized <- data$moco2[1:(length(trend))] - trend

plot(deseasonalized,
     main = "Deseasonalized Atmospheric CO2 at Mauna Loa Observatory",
     xlab = 'Year',
     ylab = 'Distance from Trend')
```

# 3.1.2

PROBLEM STATEMENT: 
For the MA(2) process represented by the following equation:

X(t) = a(t) + 0.2 a(t-1) - 0.48 a(t-2)

use the equation for autocorrelation for an MA(q) process to calculate by hand 
the following values:
ρ0, ρ1, ρ2, and ρ3.

## ρ0

We know that ρ0 of any autocorrelation function is 1.
ρ0 = 1

## ρ1
	
$$\frac{0.2 + (-0.2)(0.48)}{1 + (-0.2)^2+(0.48)^2}$$
$$\frac{1.04}{1.2704}$$

ρ1 = 0.819

We can check this answer by calculating it with R

```{r}
library('tswge')

ts <- plotts.true.wge(100, phi=c(0), theta=c(-0.2, 0.48))
autocorrelations <- ts$aut1
autocorrelations[2]

```

## ρ2

$$\frac{-0.48}{1 + (-0.2)^2+(0.48)^2}$$
$$\frac{-0.48}{1.2704}$$

ρ2 = -0.378

Check our answer again:

```{r}
autocorrelations[3]
```


## ρ3

We know that if k > q, ρk = 0. Here, k = 3 and q = 2, therefore ρ3 = 0. 

# 3.1.3
PROBLEM STATEMENT:
Calculate the autocorrelation at lags 0 through 10 for an AR(1) process with
φ1 = -0.7.

```{r}
ar.ts1 <- plotts.true.wge(100, phi=c(-0.7), theta=c(0), lag.max = 10)
ar.ts1$aut1
```





















